New insights into the Weddell Sea ecosystem applying a quantitative network approach

### 2022-12-29 Manuscript/Manuscript.Rmd
Updated biblio

<!-- Relevance of studying food webs -->
Food web analysis constitutes an important framework for understanding ecological community structure and for conserving biodiversity through ecosystem management [@Thompson2012]. Although topological food web analysis, which considers only the presence and absence of predator-prey interactions, provides important insights into the structure and functioning of ecological communities [e.g. @Pascual2006; @Kortsch2015; @Marina2018; @Cordone2020; @Rodriguez2022], more information on the nature of the trophic interactions is needed to effectively charaterise ecosystem dynamics and stability [@Kortsch2021, @otherRefs]. This is a fundamental step for providing assemessments on ecosystem vulnerability to environmental pressures and for prioritising management actions. In this regard, quantifying the strength of trohic interactions and species' roles within the network are of paramount importance [@Carrara2015; @Allesina2015; @Nilsson2016; @Cirtwill2018a].

<!-- Paragraph on unweighted and weighted networks  -->
Estimating interaction strength in food webs allows differentiating the importance of species interactions. On the contrary, unweighetd food web representations give equal importantance to all interactions well-knowing that some species interactions are stronger than others and hence play a diffrent role for ecosystem functioning and stability. Both empirical and theoretical studies show that interactions strength distributions in food webs are assymetric [@Paine1992; @McCann1998; @Emmerson2004; @Wootton2005; @Kortsch2021], containing a few strong and many weak interactions. This asymmetric patterning of weak and strong links is crucial to food web stability [@Paine1992; @McCann1998; @Neutel2022]. In a recent paper on an aquatic food web it was further highlighted that temporal changes in ecosytem functioning could only be predicted using weighted food web structure [@Kortsch2021]. Hence, in order to assess the stability and functioning of a food web, it is important to first determine the quantitative structure of the trophic network.

<!-- LAS I think that the methodology of estimating the IS is in the heart of the paper so we need to say something about that, this is why I added the following paragraph-->
Several methodologies have been applied to estimate interaction strength in food webs, where the quantity and quality of the data mostly determines which approach is the most convenient [@Berlow2004]. Approaches include experimental methods combined with dynamic modelling [@Emmerson2004; @Carrara2015], measurements of species abundances through time [@Fahimipour2014; @Chang2021], and estimation of metabolic rates and biomass of all species in the community [@Neutel2014]. However, these types of methods require large experimental set-ups and parameterisations restricting the analyses to smaller networks (e.g., approximatley 10 species or less). Other methods based on allometric scaling relationships and biomass information [@Kortsch 2021; @Gauzens2018] can be applied to larger networks with less data requirements, but this comes at the expense of precision in the predictions. For even larger food webs composed of nearly 1000 species and more than 10000 interactions, only methods with even less data requirements are feasible. One of these methods, proposed by @Pawar2012, combines data on consumer and resource body masses and consumer search space (interaction dimensionality) to obtain interaction strength estimates for each pairwise predator-prey interaction. An advantage of this method is that it can be applied without information on species biomass.[WHAT ARE THE IMPLICATION OF THIS; NOT INCLUSDING BIOMASS; WHAT IS THE ASSUMPTION WE ARE MAKING; MAYBE WE NEED TO ADD A FEW WORDS ABOUT THIS IN THE DISCUSSION]

[Leo; what do you mean with less powerful? in terms of what be explicit, do you mean precise?]

<!-- Insights of network science to polar ecosystems -->
Using a a network approach, different types (e.g., terrestial, lake, marine) of food webs from various geographic locations are studied worldwide, including marine polar food webs [@Carscallen2012; @Santana2013; @Kortsch2019; @Pecuchet2022]. Some of the studies from the Arctic show how food web properties (e.g., connectance) are constrained by environmental factors such as sea ice cover and seawater temperature [@Kortsch2019; @Pecuchet2022]. In the Southern Ocean, important insights have been gained into mechanisms of energy flow, the relative importance of individual species and their traits, and the influence of environmental variables (e.g., sea-ice) on the structure of local food webs [@Cordone2020; @Rossi2019]. For instance, in Potter Cove (West Antarctic Peninsula) the substratum type (i.e. hard/soft or rocks/sediments) plays a significant role in the structure and stability of the food web. In Terra Nova Bay (Ross Sea), the architecture of biodiversity was reshaped by the pulsed input of sympagic food sources following sea-ice break up, with food web simplification, decreased intraguild predation, potential disturbance propagation and increased vulnerability to biodiversity loss [@Rossi2019].
> [susanne] what do you mean here: the relative importance and traits of individual species? What important mechanisms? Can we be more specefic? Can we also be more specific about how environmental variables (e.g., sea-ice) influence the structure of local food webs in the SO?
> 
>  [Tomás] I added 2 phrases in this regard.

@@

> @learsar could you add this reference to Zotero? Fahrbach, E., Beszczynska-Möller, A., & Rohardt, G. (2009). Polar oceans-an oceanographic overview. In Biological Studies in Polar Oceans-Exploration of Life in Icy Waters, G. and H. Hempel (Eds.), Wirtschaftsverlag NW, Bremerhaven (pp. 17-36). https://epic.awi.de/id/eprint/23118/

@@

To identify the species with the highest effect on food web stability, we plotted the results of each species' extinction and its effect on food web stability, considering weighted (interaction strength) and unweighted properties (WHICH?), and species habitat. [THIS IS NOT 100% CLEAR; I NEED TO SEE THE FIGURES TO IMPROVE HERE, DON'T SEE THEM IN THIS FILE]

@@

Our analyses show that the distribution of species interaction strength at the network level is asymmetric, i.e., the Weddell Sea food web contains many weak interactions and only a few strong ones. This finding is consistent with many previous theoretical and empirical studies [e.g. @McCann1998; @Neutel2002; @Emmerson2004; @Wootton2005; @Kortsch2021]. The asymmetric distribution of interaction strength in food webs has been interpreted as an explanation for the persistence of complex communities in nature [@Bascompte2005; @Allesina2015; @Nilsson2016]. Here we show that this pattern is also prevalent in the Weddell Sea, one of the most complex food webs to date, comprising 490 species and 16041 predator-prey interactions. This finding is in someway validating the method we used, validation of allometric methods of interaction strength estimation not including interaction dimensionality has been performed for microcosmos of relatively few species [@Jonsson2018].

@@

We employed a range of descriptors using both unweighted and weighted food web properties to characterise the dynamic and multifaceted nature of the Weddell Sea food web. Our results show a positive relationship between interaction strength and trophic level, and between interaction strength and species degree. In the Weddell Sea, species with high degree also tend to have high mean interaction strengths. This positive relationship between interaction strength and species degree reinforces the central role of species with many interactions: species with a high degree (hubs) have a large impact on overall food web structure and functioning [@Dunne2002a; @Kortsch2015]. On the other hand, the positive interaction strength-trophic level relationship contradicts studies that suggest that mid-trophic level species (e.g. krill, mesopelagic fish, squid) are involved in the major pathways of energy flow in high-latitude marine ecosystems [@Pinkerton2014; @Murphy2016; @McCormack2020; @Riccialdelli2020]. Such contradiction could be explained by the lack of species biomass information in the calculation of interaction strength we applied here [@Pawar2012]. Although this methodology allows information on species biomass or density to be included, this type of data was not available for the majority of species of the Weddell Sea food web.

@@

Our study goes beyond the current understanding of how species influence ecosystem structure and stability in the Weddell Sea in particular, and in most polar regions in general [@Murphy2016; @McCormack2021]. In the same analysis we integrated information on weighted (interaction strength) and unweighted species properties, enabling a more complete assessment of the species’ role realted food web structure and stability, and allowing us to identify species and their characteristics which can have a destabilising or stabilising effect on the food web.

>>>>

<!-- Relevance of studying food webs -->
Food web analysis constitutes an important framework for understanding ecological community structure and for conserving biodiversity through ecosystem management [@Thompson2012]. Although topological food web analysis, which considers only the presence and absence of predator-prey interactions, provides important insights into the structure and functioning of ecological communities [e.g. @Pascual2006; @Kortsch2015; @Marina2018; @Cordone2020; @Rodriguez2022], more information on the nature of the trophic interactions is needed to effectively charaterise ecosystem dynamics and stability [@Kortsch2021]. This is a fundamental step for providing assemessments on ecosystem vulnerability to environmental pressures and for prioritising management actions. In this regard, quantifying the strength of trohic interactions and species' roles within the network are of paramount importance [@Carrara2015; @Allesina2015; @Nilsson2016; @Cirtwill2018a].

<!-- Paragraph on unweighted and weighted networks  -->
Estimating interaction strength in food webs allows differentiating the importance of species interactions. On the contrary, unweighetd food web representations give equal importantance to all interactions well-knowing that some species interactions are stronger than others and hence play a diffrent role for ecosystem functioning and stability. Both empirical and theoretical studies show that interactions strength distributions in food webs are assymetric [@Paine1992; @McCann1998; @Emmerson2004; @Wootton2005; @Kortsch2021], containing a few strong and many weak interactions. This asymmetric patterning of weak and strong links is crucial to food web stability [@Paine1992; @McCann1998; @Neutel2002]. In a recent paper on an aquatic food web it was further highlighted that temporal changes in ecosytem functioning could only be predicted using weighted food web structure [@Kortsch2021]. Hence, in order to assess the stability and functioning of a food web, it is important to first determine the quantitative structure of the trophic network.

<!-- LAS I think that the methodology of estimating the IS is in the heart of the paper so we need to say something about that, this is why I added the following paragraph-->
Several methodologies have been applied to estimate interaction strength in food webs, where the quantity and quality of the data mostly determines which approach is the most convenient [@Berlow2004]. Approaches include experimental methods combined with dynamic modelling [@Emmerson2004; @Carrara2015], measurements of species abundances through time [@Fahimipour2014; @Chang2021], and estimation of metabolic rates and biomass of all species in the community [@Neutel2014]. However, these types of methods require large experimental set-ups and parameterisations restricting the analyses to smaller networks (e.g., approximatley 10 species or less). Other methods based on allometric scaling relationships and biomass information [@Kortsch2021; @Gauzens2019] can be applied to larger networks with less data requirements, but this comes at the expense of precision in the predictions. For even larger food webs composed of nearly 1000 species and more than 10000 interactions, only methods with even less data requirements are feasible. One of these methods, proposed by @Pawar2012, combines data on consumer and resource body masses and consumer search space (interaction dimensionality) to obtain interaction strength estimates for each pairwise predator-prey interaction. An advantage of this method is that it can be applied without information on species biomass.

<!-- Insights of network science to polar ecosystems -->
Using a a network approach, different types (e.g., terrestial, lake, marine) of food webs from various geographic locations are studied worldwide, including marine polar food webs [@Carscallen2012; @Santana2013; @Kortsch2019; @Pecuchet2022]. Some of the studies from the Arctic show how food web properties (e.g., connectance) are constrained by environmental factors such as sea ice cover and seawater temperature [@Kortsch2019; @Pecuchet2022]. In the Southern Ocean, important insights have been gained into mechanisms of energy flow, the relative importance of individual species and their traits, and the influence of environmental variables (e.g., sea-ice) on the structure of local food webs [@Cordone2020; @Rossi2019]. For instance, in Potter Cove (West Antarctic Peninsula) the substratum type (i.e. hard/soft or rocks/sediments) plays a significant role in the structure and stability of the food web. In Terra Nova Bay (Ross Sea), the architecture of biodiversity was reshaped by the pulsed input of sympagic food sources following sea-ice break up, with food web simplification, decreased intraguild predation, potential disturbance propagation and increased vulnerability to biodiversity loss [@Rossi2019].

@@

@@

To identify the species with the highest effect on food web stability, we plotted the results of each species' extinction and its effect on food web stability, considering weighted (interaction strength) and unweighted properties, and species habitat.

@@

Our analyses show that the distribution of species interaction strength (IS) at the network level is asymmetric, i.e., the Weddell Sea food web contains many weak interactions and only a few strong ones. This finding is consistent with many previous theoretical and empirical studies [e.g. @McCann1998; @Neutel2002; @Emmerson2004; @Wootton2005; @Kortsch2021]. The asymmetric distribution of IS in food webs has been interpreted as an explanation for the persistence of complex communities in nature [@Bascompte2005; @Allesina2015; @Nilsson2016]. Here we show that this pattern is also prevalent in the Weddell Sea, one of the most complex food webs to date, comprising 490 species and 16041 predator-prey interactions. This finding is in someway validating the method we used, validation of allometric methods of IS estimation not including interaction dimensionality has been performed for microcosmos of relatively few species [@Jonsson2018].

@@

We employed a range of descriptors using both unweighted and weighted food web properties to characterise the dynamic and multifaceted nature of the Weddell Sea food web. Our results show a positive relationship between IS and trophic level, and between IS and species degree. In the Weddell Sea, species with high degree also tend to have high mean ISs. This positive relationship between IS and species degree reinforces the central role of species with many interactions: species with a high degree (hubs) have a large impact on overall food web structure and functioning [@Dunne2002a; @Kortsch2015]. On the other hand, the positive IS-trophic level relationship contradicts studies that suggest that mid-trophic level species (e.g. krill, mesopelagic fish, squid) are involved in the major pathways of energy flow in high-latitude marine ecosystems [@Pinkerton2014; @Murphy2016; @McCormack2020; @Riccialdelli2020]. Such contradiction could be explained by the lack of species biomass information in the calculation of IS we applied here [@Pawar2012]. Although this methodology allows information on species biomass or density to be included, this type of data was not available for the majority of species of the Weddell Sea food web.

@@

Our study goes beyond the current understanding of how species influence ecosystem structure and stability in the Weddell Sea in particular, and in most polar regions in general [@Murphy2016; @McCormack2021]. In the same analysis we integrated information on weighted (IS) and unweighted species properties, enabling a more complete assessment of the species’ role realted food web structure and stability, and allowing us to identify species and their characteristics which can have a destabilising or stabilising effect on the food web.
